MIGRACE DAT - DEMUS / MUSEION

### 2023-02-13 MSK_MUZBE_bioslovniky.Rmd
GEN_slovniky, MUZBE_slovniky

---
title: "MIGRACE DAT - DEMUS / MUSEION"
subtitle: "Moravskoslezsky kraj - Muzeum Beskyd"
author: "Katerina Krizova"
date: "2023-02-08"
output:
  pdf_document
    toc: true
    toc_depth: 2
    number_sections: false
    fig_caption: true
  fontsize: 11pt
  geometry: margin=1in

@@

# install.packages("tidyverse", dependencies = T)
# install.packages("readxl")
# install.packages("stringr")
# install.packages("mdbr")
# install.packages("RODBC")
# install.packages("odbc")
require(tidyverse)
require(readxl)
require(stringr) # ::str_split_fixed
require(mdbr) # ::read_mdb
require(RODBC) # ::read_mdb
require(odbc) # ::read_mdb

@@

# sbirka <- 
# podsbirka <-

@@

path_mdb <-  "M:/03 klienti/kraj moravskoslezsky/muzeum beskyd - MUZBE/01 muzBE analyza/"

@@

dbname <- paste0(path_mdb, "PVO/Databaze_PVO_2021-11-03.accdb") # path to database
con <- RODBC::odbcConnectAccess2007(dbname, pwd = "Fazole")      # set the connection

@@

 "ID"        "druh_CZ"   "druh_LAT"  "autor"     "celed_CZ"  "celed_LAT" "rad_CZ"    "rad_LAT"  
"trida_CZ"  "trida_LAT" "kmen_LAT"  "kmen_CZ"   "ochrana"   "syn"

@@

animalia <- db_tab %>% 
   separate(celed,into=c("celed_CZ", "celed_LAT") , sep=" ") %>% 
   separate(rad,into=c("rad_CZ", "rad_LAT") , sep=" ") %>% 
   separate(trida,into=c("trida_CZ", "trida_LAT") , sep=" ") %>% 
   separate(kmen,into=c("kmen_LAT", "kmen_CZ") , sep=" - ") %>% 
   mutate(celed_LAT = gsub("[()]", "", celed_LAT),
          rad_LAT = gsub("[()]", "", rad_LAT), 
          trida_LAT = gsub("[()]", "", trida_LAT)) %>% 
  rename(druh_LAT = druh_L) %>% 
  rename(druh_CZ = druh_cz)

@@

>>>>

---
title: "Moravskoslezsky kraj - Muzeum Beskyd"
subtitle: "Tvorba a kontrola bioslovníků"
author: "Katerina Krizova"
date: "2023-02-08"
output:
  pdf_document
    toc: true
    toc_depth: 2
    number_sections: false
    fig_caption: true
  fontsize: 11pt
  geometry: margin=1in

@@

# install.packages("tidyverse", dependencies = T)
# install.packages("readxl")
# install.packages("stringr")
# install.packages("mdbr")
# install.packages("RODBC")
# install.packages("odbc")
# install.packages("xlsx")
require(tidyverse)
require(readxl)
require(stringr) # ::str_split_fixed
require(mdbr) # ::read_mdb
require(RODBC) # ::read_mdb
require(odbc) # ::read_mdb
require(xlsx) # write excel sheets

@@

# slovnik <- "animalia"
slovnik <- "fungi"
# slovnik <- "plantae"

@@

path_mdb <-  "M:/03 klienti/kraj moravskoslezsky/muzeum beskyd - MUZBE/01 muzBE analyza/PVO/"

# NACTENI BIOLIB SLOVNÍKU

Kmen  
Třída  
Řád  
Čeleď  
Rod  
Druh  
Poddruh

```{r LOAD BIOLIB SUMMARY KK}

biolib <- read.table(paste0(path, "bioslov/BioLib_KKsummary_", slovnik, ".csv"), 
                     sep = ";", header = T)

# biolib[is.na(biolib)] <- ""  # replace NA

# check Nomen and NadNomen categories
# ? are they the same or do they differ?
# ? which column to keep?

biolib$diff = ifelse(biolib$kmen_Nomen == biolib$trida_NadNomen, 'stejne', 'lisi se') # keep 'kmen'
biolib$diff = ifelse(biolib$trida_Nomen == biolib$rad_NadNomen, 'stejne', 'lisi se') # keep 'rad_NadNomen'
biolib$diff = ifelse(biolib$rad_Nomen == biolib$celed_NadNomen, 'stejne', 'lisi se') # keep 'celed_NadNomen'
biolib$diff = ifelse(biolib$celed_Nomen == biolib$rod_NadNomen, 'stejne', 'lisi se') # keep 'rod_nadNomen'
biolib$diff = ifelse(biolib$rod_Nomen == biolib$druh_NadNomen, 'stejne', 'lisi se') # keep 'rod_Nomen'
biolib$diff = ifelse(biolib$druh_Nomen == biolib$poddruh_NadNomen, 'stejne', 'lisi se') # keep 'druh_Nomen'

unique(biolib$diff)

# select columns based on above

sel_biol <- biolib %>% 
  select(kmen_Nomen, kmen_Autor,             # kmen
         rad_NadNomen, trida_Autor,          # trida
         celed_NadNomen, rad_Autor,          # rad
         rod_NadNomen, celed_Autor,          # celed    
         rod_Nomen, rod_Autor,               # rod
         druh_Nomen, druh_Autor              # druh
         # , poddruh_Nomen, poddruh_Autor    # poddruh (only if present in data)
         )  %>% 
 unite(sys_tree_BIOL, c("kmen_Nomen", "rad_NadNomen", "celed_NadNomen", 
                   "rod_NadNomen", "rod_Nomen", "druh_Nomen"
                   # , "poddruh_Nomen"       # poddruh (only if present in data)
                   ), 
       sep = "/", na.rm = T, remove = FALSE) %>% 
  select(sys_tree_BIOL, druh_Nomen, druh_Autor) %>% 
  distinct()

@@

dbname <- paste0(path_mdb, "Databaze_PVO_2021-11-03.accdb") # path to database
con <- RODBC::odbcConnectAccess2007(dbname, pwd = "Fazole")      # set the connection

@@

@@

animalia <- db_tab %>% 
   separate(celed,into=c("celed_CZ", "celed_LAT") , sep=" ") %>% 
   separate(rad,into=c("rad_CZ", "rad_LAT") , sep=" ") %>% 
   separate(trida,into=c("trida_CZ", "trida_LAT") , sep=" ") %>% 
   separate(kmen,into=c("kmen_LAT", "kmen_CZ") , sep=" - ") %>% 
   mutate(celed_LAT = gsub("[()]", "", celed_LAT),
          rad_LAT = gsub("[()]", "", rad_LAT), 
          trida_LAT = gsub("[()]", "", trida_LAT),
          rod_LAT = gsub(" .*$", "", druh_L)) %>% 
  rename(druh_LAT = druh_L) %>% 
  rename(druh_CZ = druh_cz)

@@

sel_anim <- animalia %>% 
  select(kmen_LAT, trida_LAT, rad_LAT, celed_LAT, rod_LAT, druh_LAT, autor) %>% 
   unite(sys_tree_ANIM, c("kmen_LAT", "trida_LAT", "rad_LAT", 
                   "celed_LAT", "rod_LAT", "druh_LAT"
                   # , "poddruh_LAT"       # poddruh (only if present in data)
                   ), 
       sep = "/", na.rm = T, remove = FALSE) %>% 
  select(sys_tree_ANIM, druh_LAT, autor) %>% 
  distinct()

### porovnani

```{r COMPARE ANIMALIA W BIOLIB}

comp <- left_join(sel_anim, sel_biol, by = c("druh_LAT" = "druh_Nomen"), keep = T, multiple = "any") %>% 
  filter(sys_tree_ANIM != sys_tree_BIOL)

# systematic tree diffs

diff_tree <- comp
diff_tree$diff = ifelse(diff_tree$sys_tree_ANIM == diff_tree$sys_tree_BIOL, 'stejne', 'lisi se')

diff_tree <- diff_tree %>% 
  rename("STROM_data" = sys_tree_ANIM, "STROM_biolib" = sys_tree_BIOL,
         "AUTOR_data" = autor, "AUTOR_biolib" = druh_Autor)

# author diffs

diff_author <- comp %>% 
  mutate(autor = str_replace_all(autor, " ", "")) %>% 
  mutate(druh_Autor = str_replace_all(druh_Autor, " ", ""))

diff_author$diff = ifelse(diff_author$autor == diff_author$druh_Autor, 'stejne', 'lisi se')

diff_author <- diff_author %>%  
 filter(diff == "lisi se") %>% 
  rename("STROM_data" = sys_tree_ANIM, "STROM_biolib" = sys_tree_BIOL,
         "AUTOR_data" = autor, "AUTOR_biolib" = druh_Autor)

# write excel with differences

library(xlsx)
write.xlsx(diff_tree, file=paste0(path_mdb, project, "_bioslovniky_zoo_ROZDILY.xlsx"), 
           sheetName="systematika", row.names=FALSE)
write.xlsx(diff_author, file=paste0(path_mdb, project, "_bioslovniky_zoo_ROZDILY.xlsx"), 
           sheetName="autori", row.names=FALSE, append=TRUE)

### 2023-02-13 _GEN_BIOLIB_slovniky.Rmd
GEN_slovniky, MUZBE_slovniky

project <- "MSK_MUZBE"

# sbirka <- 
# podsbirka <-

@@

path_data <-  paste0(path, "data/")
path_csv <-  paste0(path, "mus_csv/", project, "/")

path_mdb <-  "M:/03 klienti/kraj moravskoslezsky/muzeum beskyd - MUZBE/"

# NACTENI BIOLIB SLOVNIKU (KK sum)

```{r BIOLIB DICT}

biolib <- read.table(paste0(path, "BioLib_KKsummary.csv"), sep = ";", header = T)

@@

dbname <- paste0(path_mdb, "01 muzBE analyza/PVO/Databaze_PVO_2021-11-03.accdb") # path to database
con <- RODBC::odbcConnectAccess2007(dbname, pwd = "Fazole")      # set the connection

RODBC::sqlTables(con)                                            # list tables in database

@@

## ANIMALIA

```{r READ AND MODIFY DB TABLES FOR ANIMALIA, warning = F}

# read all taxons

zoo <- RODBC::sqlFetch(con, "Pz_System_ZOOpredmety")

## FUNGI

```{r READ AND MODIFY DB TABLE FUNGHI, warning = F}

kmen <- RODBC::sqlFetch(con, "PRM_BioLib_080_Fungi_Kmen")  
trida <- RODBC::sqlFetch(con, "PRM_BioLib_090_Fungi_Trida")  
rad <- RODBC::sqlFetch(con, "PRM_BioLib_100_Fungi_Rad")  
celed <- RODBC::sqlFetch(con, "PRM_BioLib_110_Fungi_Celed")  
rod <- RODBC::sqlFetch(con, "PRM_BioLib_130_Fungi_Rod")  
druh <- RODBC::sqlFetch(con, "PRM_BioLib_150_Fungi_Druh")  
poddruh <- RODBC::sqlFetch(con, "PRM_BioLib_153_Fungi_Poddruh")

## PLANTAE

```{r}

kmen <- RODBC::sqlFetch(con, "PRM_BioLib_080_Plantae_Kmen")  
trida <- RODBC::sqlFetch(con, "PRM_BioLib_090_Plantae_Trida")  
rad <- RODBC::sqlFetch(con, "PRM_BioLib_100_Plantae_Rad")  
celed <- RODBC::sqlFetch(con, "PRM_BioLib_110_Plantae_Celed")  
rod <- RODBC::sqlFetch(con, "PRM_BioLib_130_Plantae_Rod")  
druh <- RODBC::sqlFetch(con, "PRM_BioLib_150_Plantae_Druh")  
poddruh <- RODBC::sqlFetch(con, "PRM_BioLib_153_Plantae_Poddruh")

@@

@@

# full_join - this setup gives full table WITH ALL and ORDERED COLUMNS

write.table(join6, paste0(path, "BioLib_KKsummary.csv"), sep = ";", row.names = F)

#   p  i  s  k  o  v  i  s  t  e  -  ZACATEK

@@

#   p  i  s  k  o  v  i  s  t  e  -  KONEC

# remove prelim dfs

rm(ani1)
rm(ani2)
rm(ani3)
rm(ani4)
rm(ani5)

>>>>

# slovnik <- "animalia"
slovnik <- "fungi"
# slovnik <- "plantae"

@@

path_mdb <-  "M:/02 dokumenty/03 migrace/BIO/"

@@

dbname <- paste0(path_mdb, "bio_slovniky.accdb") # path to database
con <- RODBC::odbcConnectAccess2007(dbname)      # set the connection

# RODBC::sqlTables(con)                           # list tables in database

@@

```{r READ AND MODIFY DB TABLES, warning = F}

if(slovnik == "animalia"){
  kmen <- RODBC::sqlFetch(con, "PRM_BioLib_080_Animalia_Kmen")  
  trida <- RODBC::sqlFetch(con, "PRM_BioLib_090_Animalia_Trida")  
  rad <- RODBC::sqlFetch(con, "PRM_BioLib_100_Animalia_Rad")  
  celed <- RODBC::sqlFetch(con, "PRM_BioLib_110_Animalia_Celed")  
  rod <- RODBC::sqlFetch(con, "PRM_BioLib_130_Animalia_Rod")  
  druh <- RODBC::sqlFetch(con, "PRM_BioLib_150_Animalia_Druh")  
  poddruh <- RODBC::sqlFetch(con, "PRM_BioLib_153_Animalia_Poddruh") 
} else if(slovnik == "fungi"){
  kmen <- RODBC::sqlFetch(con, "PRM_BioLib_080_Fungi_Kmen")  
  trida <- RODBC::sqlFetch(con, "PRM_BioLib_090_Fungi_Trida")  
  rad <- RODBC::sqlFetch(con, "PRM_BioLib_100_Fungi_Rad")  
  celed <- RODBC::sqlFetch(con, "PRM_BioLib_110_Fungi_Celed")  
  rod <- RODBC::sqlFetch(con, "PRM_BioLib_130_Fungi_Rod")  
  druh <- RODBC::sqlFetch(con, "PRM_BioLib_150_Fungi_Druh")  
  poddruh <- RODBC::sqlFetch(con, "PRM_BioLib_153_Fungi_Poddruh")
} else if(slovnik == "plantae"){
  kmen <- RODBC::sqlFetch(con, "PRM_BioLib_080_Plantae_Kmen")  
  trida <- RODBC::sqlFetch(con, "PRM_BioLib_090_Plantae_Trida")  
  rad <- RODBC::sqlFetch(con, "PRM_BioLib_100_Plantae_Rad")  
  celed <- RODBC::sqlFetch(con, "PRM_BioLib_110_Plantae_Celed")  
  rod <- RODBC::sqlFetch(con, "PRM_BioLib_130_Plantae_Rod")  
  druh <- RODBC::sqlFetch(con, "PRM_BioLib_150_Plantae_Druh")  
  poddruh <- RODBC::sqlFetch(con, "PRM_BioLib_153_Plantae_Poddruh")
} else print("I have a bad feeling about this!")

@@

# full_join - this setup gives a full table WITH ALL and ORDERED COLUMNS

write.table(join6, paste0(path, "BioLib_KKsummary_", slovnik, ".csv"), sep = ";", row.names = F)

# p i s k o v i s t e

```{r sandbox, eval = F}

@@

@@
